---
title: "QA"
author: "Will Ball"
date: "24/08/2021"
output: html_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)

.libPaths("H:/DaSH477/Methods/R Libraries")
```

# Packages

```{r packages, message=FALSE,warning=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(vroom)
library(skimr)
library(finalfit)
library(readr)
library(tidylog)
library(data.table)

# pis_original <- vroom("H:/DaSH477/Data/Cleaned Data/Will/original_pis.csv", delim = ",")
```

# Data

Pipe ('|') is delim

pis_original <- vroom("H:/DaSH477/Data/Cleaned Data/Will/original_pis.csv", delim = ",")

```{r data, message=FALSE,warning=FALSE}
# 1. Import original data. Rename and derive
pis_original <- vroom("H:/DaSH477/Data/Raw Data/Stage3_20211108/DaSH477_PIS_Release_v3.txt", delim = "|") %>% 
  clean_names() %>%
#  slice(-(1)) %>%  # Remove row of dashes
#  mutate_if(is.character, ~ifelse(. == "NULL", NA, .)) %>% 
  rename(dash_uid = da_sh477_study_num,                            # rename for clarity
         datazone_2012 = dz_2012,
         datazone_2016 = dz_2016,
         datazone_2020 = dz_2020v2,
         bnf_item = bnf_item_code) %>%
  mutate(bnf_chapter = str_sub(bnf_item, start = 1, end = 2),      # Derive BNF chapter/section/subsection/paragraph codes
         bnf_section = str_sub(bnf_item, start = 1, end = 4),
         bnf_subsec = str_sub(bnf_item, start = 1, end = 6),
         bnf_para = str_sub(bnf_item, start = 1, end = 7),
         month_of_birth = as_date(month_of_birth),                     # Change from datetime class to just date
         chi_mob = as_date(chi_mob),
         pis_date = as_date(pis_date),
         paid_date = as_date(paid_date),
         prescribed_date = as_date(prescribed_date),
         dispensed_date = as_date(dispensed_date))

##### Replace NULL values with NA


# 2. Derive age at event & age group
pis_original <- pis_original %>% 
  mutate(age_at_event = (month_of_birth %--% pis_date %/% years(1)),
         age_group = case_when(
                      age_at_event %in% c(0:4)   ~ "0-4",
                      age_at_event %in% c(5:9)   ~ "5-9",
                      age_at_event %in% c(10:14) ~ "10-14",
                      age_at_event %in% c(15:19) ~ "15-19",
                      age_at_event %in% c(20:24) ~ "20-24"),
         age_group = ordered(age_group, levels = c("0-4","5-9","10-14","15-19","20-24")))

# 3. Save Original
fwrite(pis_original, "H:/DaSH477/Data/Cleaned Data/Will/original_pis.csv", sep = ",")

```

# Missingness & Checks

```{r NAs, message=FALSE,warning=FALSE}
# 0. Skim
# skim_pis <- skim(pis_original)

# 1. Missing-ness
# missing_plot(pis_original)
missing_pis <- missing_glimpse(pis_original) %>% 
  arrange(desc(missing_n))
missing_pis

# 2. Checks
checks <- pis_original %>%
  summarise(
    rows = nrow(pis_original),
    distinct = n_distinct(pis_original),
    individuals = n_distinct(dash_uid),
    date_min = min(prescribed_date, na.rm = TRUE),
    date_max = max(prescribed_date, na.rm = TRUE),
    age_min = min(age_at_event, na.rm = TRUE),
    age_max = max(age_at_event, na.rm = TRUE)
  )
checks
```

# Age Checks

Some records with negative ages. (Fixable)
Some records with NA month_of_birth. (Fixable)
Some where age_at_event >= 25. (?Remove)

```{r ages, warning=F,message=F}
# 3. Ages
pis_original %>% 
  count(age_at_event)                      # Negative ages & some people aged 25 at event

minus_age <- pis_original %>%
  filter(age_at_event < 0)                 # Dataset mob is after each pis_date. chi_mob seems more reasonable for this case

# 4. NA ages and age groups
na_age <- pis_original %>% 
  filter(is.na(age_at_event))

na_age %>% 
  count(month_of_birth, dob_source)        # Source is CHI, but no month_of_birth recorded

na_age_group <- pis_original %>% 
  filter(is.na(age_group))

na_age_group %>% 
  count(age_at_event)

# 5. Aged 25
age_25 <- pis_original %>% 
  filter(age_at_event == 25) %>% 
  select(dash_uid, pis_date, month_of_birth, chi_mob, dob_source, age_at_event, age_group) %>%
  mutate(days = (month_of_birth %--% pis_date %/% days(1)),
         years = days/365.25,
         mob_agree = case_when(
           month_of_birth == chi_mob ~ 1,
           TRUE                      ~ 0)
         )

age_25 %>% 
  count(dob_source)                        # All from dataset, none from CHI

age_25 %>% 
  count(mob_agree)                         # All agree with CHI mob
```

# Items without a BNF code

Lots of DUMMY records.
Some named items without BNF item code. (Fixable)

```{r bnf item, warning=F,message=F}
rm(checks, missing_pis)

# 1. Check records without BNF Item Code
no_bnf_item <- pis_original %>%
  filter(is.na(bnf_item))

no_bnf_item %>%
  count(approved_name, sort = TRUE)     # Nearly all 'DUMMY'. Can complete BNF chapter/section/subsec/para

# x. Dummy only
dummy <- pis_original %>% 
  filter(approved_name == "DUMMY")      # According to ISD documentation, dummy items weren't recognised or payment rejected

dummy %>%
  count(prescribable_item_name, sort = T)
```

# Univariable plots

Date - Span seems reasonable. Trend reducing, 2x drops around 2019 & 2021
Age at event - Expected trend
Sex - No NAs, expected proportions
Age Group - NAs (See above)

```{r explore, warning=F,message=F}
qplot(x = year(pis_date), data = pis_original, geom = "histogram")

qplot(x = age_at_event, data = pis_original, geom = "histogram", stat = "bin", binwidth = 1)

qplot(x = age_group, data = pis_original)

qplot(x = chi_sex, data = pis_original)

qplot(x = bnf_chapter, data = pis_original)
```

# Top 20s

```{r top 20s, warning=F,message=F}
pis_original %>%                       # All records
  count(approved_name, sort = T) %>%
  top_n(20) %>% 
  mutate(rank = row_number()) %>% 
  select(rank, everything())


pis_original %>% 
  filter(bnf_chapter == "04") %>%       # CNS Chapter
  count(approved_name, sort = T) %>%
  top_n(20) %>% 
  mutate(rank = row_number()) %>% 
  select(rank, everything())
```
