---
title: "PIS Cleaning"
author: "Will Ball"
date: "25/08/2021"
output: html_document
---
# negative ages handling modified by Helen Rowlands 27/04/2022 


# To Do:

## Age
1. NA for month_of_birth. Use chi_mob instead

2. Negative age_at_event. Use chi_mob instead of month_of_birth

3. People aged 25. Keep records as they are likely to be rounding errors (from month of birth)

## BNF Item Codes
1. Some without BNF Item Code, but approved_name can be used to assign chapter, section, subsec & paragraph codes


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)

.libPaths("H:/DaSH477/Methods/R Libraries")
```

# Packages

```{r packages, message=FALSE,warning=FALSE}
library(tidyverse)
library(lubridate)
library(janitor) 
library(vroom)
library(skimr)
library(finalfit)
library(viridis)
library(data.table)
library(tidylog)
```

# Data

```{r data, warning=FALSE,message=FALSE}
pis_original <- vroom("H:/DaSH477/Data/Cleaned Data/Will/original_pis.csv", delim = ",") # fread drops leading 0s by default

# For inclusion
pis_original %>% 
  summarise(
    rows = n(),
    individuals = n_distinct(dash_uid)
  )
```

# Age

## 1. NA month_of_birth

```{r age 2, warning=FALSE,message=FALSE}
pis_cleaned <- pis_original

# 1. Get records where age_at_event or month_of_birth are NA
na_age <- pis_cleaned %>% 
  filter(is.na(month_of_birth)) %>%
  select(dash_uid, month_of_birth, chi_mob, dob_source)

missing_plot(na_age) # Complete chi_mob

# 2. Explore source
na_age %>% 
  count(dob_source) # All chi_dob but showing NA.

# 3. Replace NAs with chi_mob
pis_cleaned$month_of_birth[is.na(pis_cleaned$month_of_birth)] <- pis_cleaned$chi_mob[is.na(pis_cleaned$month_of_birth)]

# 4. Check it worked
na_age_cleaned <- pis_cleaned %>%
  filter(is.na(month_of_birth))  # No more records with NA month_of_birth

# 5. Derive age at event and age groups again
pis_cleaned <- pis_cleaned %>% 
  mutate(age_at_event = (month_of_birth %--% pis_date %/% years(1)),
         age_group = case_when(
                      age_at_event %in% c(0:4)   ~ "0-4",
                      age_at_event %in% c(5:9)   ~ "5-9",
                      age_at_event %in% c(10:14) ~ "10-14",
                      age_at_event %in% c(15:19) ~ "15-19",
                      age_at_event %in% c(20:24) ~ "20-24"),
         age_group = ordered(age_group, levels = c("0-4","5-9","10-14","15-19","20-24")))

# 6. Check for ages at event
na_age_cleaned <- pis_cleaned %>%
  filter(is.na(age_at_event))
```

## 1. Negative age_at_event

Erroneous birth dates identified.
- Where chi_dob and month_of_birth (from Health Intelligence) are different, the realistic one has been used.
- Where month_of_birth and chi_mob are the same, records have been removed: All source dob from CHI (without century info - could be 19xx rather than 20xx). Some individuals whose medications suggest realistically they may be 100+

```{r age 1, warning=FALSE,message=FALSE}
rm(na_age, na_age_cleaned)

# 1. Identify records where month_of_birth is after pis_date
minus_age <- pis_cleaned %>%
  filter(month_of_birth >= pis_date) %>%
  select(dash_uid, pis_date, pis_date_source, paid_date, prescribed_date, month_of_birth, chi_mob, dob_source) %>% 
  mutate(mob_agree = case_when(
           month_of_birth == chi_mob ~ 1,
           TRUE                      ~ 0))

# 2. Check agreement of birth dates
minus_age %>%
  count(mob_agree)          # Most agree, some don't

n_distinct(minus_age$dash_uid) # number of individuals

# 3. Where dobs don't agree
minus_age_disagree <- minus_age %>% 
  filter(mob_agree == 0)

n_distinct(minus_age_disagree$dash_uid) 

## modified by HR

# 2. Check any other issues 
#pis_cleaned %>%
#  filter(dash_uid in  unique(minus_age_disagree$dash_uid))        # month_of_birth seems to be wrong. chi_mob makes more sense


ids <- unique(minus_age_disagree$dash_uid)

# 3. Replace erroneous month_of_birth with chi_mob for all records with id of negative aged individual
pis_cleaned$month_of_birth[pis_cleaned$dash_uid %in% ids[ids$dash_uid]] <- pis_cleaned$chi_mob[pis_cleaned$dash_uid %in% ids[ids$dash_uid]]

## end of modification

# re-do age at event & age_group
pis_cleaned <- pis_cleaned %>% 
  mutate(age_at_event = (month_of_birth %--% pis_date %/% years(1)),
         age_group = case_when(
                      age_at_event %in% c(0:4)   ~ "0-4",
                      age_at_event %in% c(5:9)   ~ "5-9",
                      age_at_event %in% c(10:14) ~ "10-14",
                      age_at_event %in% c(15:19) ~ "15-19",
                      age_at_event %in% c(20:24) ~ "20-24"),
         age_group = ordered(age_group, levels = c("0-4","5-9","10-14","15-19","20-24")))

# 4. Check records where month_of_birth and chi_mob agree
minus_age_agree <- minus_age %>% 
  filter(mob_agree == 1)

# 5. Visualise
minus_age_agree %>%
  mutate(dash_uid = factor(dash_uid)) %>%
  ggplot(aes(x = month_of_birth, y = pis_date, colour = dash_uid)) +
  scale_x_date(limits = c(ymd("2012-01-01"),max(minus_age$month_of_birth))) +
  geom_point(size = 2) +
  geom_abline(slope = 1, intercept = 0, size = 1.1, linetype = "dashed") +
  theme_minimal()

# 6. ?Wrong century
minus_age_agree_ids <- unique(minus_age_agree$dash_uid)

minus_age_agree_all <- pis_cleaned %>%
  filter(dash_uid %in% minus_age_agree_ids) %>%
  select(dash_uid, pis_date, pis_date_source, paid_date, prescribed_date, month_of_birth, chi_mob, dob_source, approved_name) %>%
  mutate(
    age_20 = (month_of_birth %--% pis_date %/% years(1)),
    age_19 = (month_of_birth %--% pis_date %/% years(1)) + 100) %>%
  group_by(dash_uid) %>%
  arrange(pis_date, .by_group = T) %>%
  mutate(record = row_number()) %>%
  ungroup()

# 7. Remove
pis_cleaned <- pis_cleaned %>%
  filter(!(dash_uid %in% minus_age_agree_ids))

pis_cleaned %>% 
  summarise(
    rows = n(),
    individuals = n_distinct(dash_uid)
  )
```

## 3. Age 25

```{r age 3, warning=FALSE,message=FALSE}
rm(minus_age, minus_age_agree, minus_age_disagree, minus_age_agree_ids, minus_age_agree_all)

# 1. Identify people aged 25 at event
age_25 <- pis_cleaned %>% 
  filter(age_at_event == 25) %>% 
  select(dash_uid, pis_date, month_of_birth, chi_mob, dob_source, age_at_event, age_group) %>%
  mutate(days = (month_of_birth %--% pis_date %/% days(1)),
         years = days/365.25,                                           # Crude approach
         years_lubridate = (month_of_birth %--% pis_date %/% years(1)), # Takes into account leap year days true to calendar
         mob_agree = case_when(
           month_of_birth == chi_mob ~ 1,
           TRUE                      ~ 0)
         )

# 2. Confirm
age_25 %>% 
  count(mob_agree)    # All dates agree with chi record

min(age_25$years)
max(age_25$years)
min(age_25$years_lubridate)
max(age_25$years_lubridate)

######### Decided to keep - DOB rounded to MOB for transfer to DaSH
# 3. Remove records where age >= 25
#pis_cleaned <- pis_cleaned %>% 
#  filter(!age_at_event == 25)

#min(pis_cleaned$age_at_event)
#max(pis_cleaned$age_at_event)

rm(age_25)
```

# BNF

## 1. Missing BNF Item Codes

1% of all records have no bnf_item codes. Nearly all of those are 'Dummy' records
- Likely to remove
- Small number of real drugs and some of these have multiple appearances in the BNF so difficult to replace NAs

```{r BNF 1, warning=FALSE,message=FALSE}
# pis_cleaned <- vroom("H:/DaSH477/Data/Cleaned Data/Will/cleaned_pis.csv", delim = ",")

# 1. Located records without bnf_item code
no_bnf_item <- pis_cleaned %>%
  filter(is.na(bnf_item) | bnf_item == "NULL")

no_bnf_item %>%
  count(approved_name, sort = TRUE)

# 2. Remove Dummy items
pis_cleaned <- pis_cleaned %>% 
  filter(!(approved_name == "DUMMY"))

pis_cleaned %>% 
  summarise(
    n = n(),
    individuals = n_distinct(dash_uid)
  )

# Repeat counts
no_bnf_item <- pis_cleaned %>%
  filter(is.na(bnf_item) | bnf_item == "NULL")

no_bnf_item %>%
  count(approved_name, sort = TRUE)

# 3. List of drugs without bnf_item codes
no_bnf_meds <- unique(no_bnf_item$approved_name)

# 4. Get most common bnf chapters, sections, subsections, paras for each drug above
meds1 <- pis_original %>%
  filter(approved_name %in% no_bnf_meds) %>%
  filter(!is.na(bnf_chapter)) %>% 
  count(approved_name, bnf_chapter, bnf_section, bnf_subsec, bnf_para) %>% 
  group_by(approved_name) %>%
  arrange(desc(n), .by_group = T) %>% 
  top_n(1)

# 5. If bnf_item is NA, replace chapter/section/subsection/para with codes from meds1
pis_cleaned <- pis_cleaned %>% 
  left_join(meds1, by = "approved_name") %>% 
              mutate(bnf_chapter = coalesce(bnf_chapter.x, bnf_chapter.y),
                     bnf_section = coalesce(bnf_section.x, bnf_section.y),
                     bnf_subsec = coalesce(bnf_subsec.x, bnf_subsec.y),
                     bnf_para = coalesce(bnf_para.x, bnf_para.y)) %>%
  select(!ends_with(c(".x",".y")))

pis_cleaned %>% 
  filter(approved_name %in% no_bnf_meds) %>% 
  filter(is.na(bnf_item) | bnf_item == "NULL") %>% 
  count(approved_name, bnf_chapter, bnf_section, bnf_subsec, bnf_para)

# 6. Ichthammol in yellow soft paraffin by hand
pis_cleaned$bnf_chapter[pis_cleaned$approved_name == "ICHTHAMMOL IN YELLOW SOFT PARAFFIN"] <- "13"
pis_cleaned$bnf_section[pis_cleaned$approved_name == "ICHTHAMMOL IN YELLOW SOFT PARAFFIN"] <- "1305"
pis_cleaned$bnf_subsec[pis_cleaned$approved_name == "ICHTHAMMOL IN YELLOW SOFT PARAFFIN"] <- "130501"
pis_cleaned$bnf_para[pis_cleaned$approved_name == "ICHTHAMMOL IN YELLOW SOFT PARAFFIN"] <- "1305010"
```

# Save

```{r save, warning=FALSE,message=FALSE}
rm(meds1, no_bnf_item, no_bnf_meds, pis_original)

fwrite(pis_cleaned, "H:/DaSH477/Data/Cleaned Data/Will/cleaned_pis.csv", sep = ",")
```